---
title: "PR02 Streaming Services Data"
author: "Kaitlynn, Ava, and Brittney"
date: '`r format(Sys.time(), "%A, %B %d, %Y @ %I:%M %p")`'
output: 
  html_document: 
    theme: cosmo
    highlight: textmate
    code_folding: hide
---

```{r globaloptions, include = FALSE}
knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE,
  comment = NA
)
```
***

## Setup

### Packages Used

```{r loadpackages}
library(tidyverse)
library(readr)
library(skimr)
library(ggplot2)
```

### Datasets Used

### Background of the Data

Two files containing data about thge content of streaming services were downloaded from Kaggle:  [Disney+ Movies and TV Shows](https://www.kaggle.com/datasets/shivamb/disney-movies-and-tv-shows) and [TV shows on Netflix, Prime Video, Hulu and Disney+](https://www.kaggle.com/datasets/ruchi798/tv-shows-on-netflix-prime-video-hulu-and-disney). You can read about them there, including variable definitions, sources, when they were created, and other information. Read the two datasets from the `data_raw` folder and use `skim()` to explore their structures.

#### Disney+ Content Dataset

```{r readdisneydata}
disney_data <- read.csv("~/git_repos/STT2860S23project2/data_raw/disneypluscontent_2021_11_21.csv")
```

```{r skimdisneydata}
skim(disney_data)
```

#### Streaming TV Content Dataset

```{r readstreamingdata}
streaming_data <- read_csv("~/git_repos/STT2860S23project2/data_raw/streamingtvcontent_2021.csv")
```

```{r skimstreamingdata}
skim(streaming_data)
```

***

## Analyses

### Clean, Tidy, and Save the Data

#### Dataset #1: Disney+ Data

The Disney+ dataset contains a variable that gives information about the `rating` for each program offering. Convert that variable into a factor that puts the ratings in (roughly) ascending order of age restriction/audience: `TV-Y`, `TV-Y7`, `TV-Y7-FV`, `TV-G`, `G`, `TV-PG`, `PG`, `PG-13`, `TV-14`. Delete the `show_id`, `director`, `cast`, and `description` variables. You should be able to do this all in one piped sequence.

```{r factor&select}
tidy_disney_data <- disney_data %>%
  select(-c(show_id, director, cast, description)) %>%
  mutate(rating = factor(rating, levels = c("TV-Y", "TV-Y7", "TV-Y7-FV", "TV-G", "G", "TV-PG", "PG", "PG-13", "TV-14")))
```

#### Dataset #2: Streaming TV Data

Perform the following manipulations to tidy the streaming TV data for analysis. You should be able to do this all in one piped sequence.

* Pivot the `Netflix`, `Hulu`, `Prime Video`, and `Disney+` variables into two columns called `Service` (containing the name of the streaming service) and `YesNo` (containing the 0/1 values from the original columns).
* Filter the dataset to remove any row where `YesNo == 0` (i.e., a show is *not* on the service).
* Separate the `IMDb` variable to split each show's actual rating from the maximum possible rating of `10`. 
* Convert the show's actual IMDB rating into a numerical variable and get rid of the column containing `10`.
* Separate the `Rotten Tomatoes` variable to split each show's actual rating from the maximum possible rating of `100`. 
* Convert the show's actual Rotten Tomatoes rating into a numerical variable and get rid of the column containing `100`.

```{r pivotandtidy}
tidy_streaming_data <- streaming_data %>%
  pivot_longer(cols = c("Netflix", "Hulu", "Prime Video", "Disney+"),
               names_to = "Service",
               values_to = "YesNo") %>%
  filter(YesNo == 1) %>%
  separate(IMDb, c("IMDb_score", "max_score"), sep = "/") %>%
  mutate(IMDb_score = as.numeric(IMDb_score)) %>%
  select(-max_score) %>%
  separate("Rotten Tomatoes", c("rt_score", "rt_max_score"), sep = "/") %>%
  mutate(rt_score = as.numeric(rt_score)) %>%
  select(-rt_max_score)
```

#### Save the Formatted Data Files

Save both data files in a single`.RData` file in the `data_tidy` folder.

```{r savedatafiles}
save(tidy_disney_data, tidy_streaming_data, file = "~/git_repos/STT2860S23project2/data_tidy/tidy_disney_streaming.RData")
```

### Analysis 1: Disney Content & Release Years

**This analysis uses Dataset #1.**

#### Release Years by Program Type---Plot #1

Create a violin plot of `release_year` (x axis) grouped by `type` of program (y axis) for programs on Disney+. Add a boxplot inside the violin plot. Re-scale the X axis so that tick marks appear at whole-decade intervals (e.g., 1980, 1990, 2000). Include an informative title and axis labels. Add at least two other types of formatting to enhance the effectiveness and/or appearance of your plot (e.g., themes, colors, etc).

```{r violinplot}
violin_disney <- tidy_disney_data %>%
  ggplot(aes(x = release_year, y = type)) +
  geom_violin() +
  geom_boxplot(width = 0.2, fill = "lightblue2", outlier.size = 0.5) +
  scale_x_continuous(breaks = seq(1925, 2025, by = 10)) +
  labs(title = "Movie and TV Show Release Year Distribution", 
       x = "Release Year",
       y = "Type") +
  theme_bw()
violin_disney
```

#### Release Years by Program Type---Plot #2

Create a histogram for `release_year` with `type` as a fill variable. Choose a reasonable number of bins or bin width. Re-scale the X axis so that tick marks appear at whole-decade intervals (e.g., 1980, 1990, 2000). Scale the Y axis to have reasonably spaced tick marks. Add vertical lines with annotations to indicate the debuts of the Disney Brothers Cartoon Studio (1923), the Disney Channel on TV (1983), and Disney+ streaming service (2019). Include an informative title and axis labels. Delete the title "type" from the legend. Add at least two other types of formatting to enhance the effectiveness and/or appearance of your plot (e.g., themes, colors, etc).

```{r histogram}
ggplot(tidy_disney_data, aes(x = release_year, fill = type)) +
  geom_histogram(bins = 20, alpha = 0.7) +
  scale_x_continuous(breaks = seq(1925, 2025, by = 10)) +
  scale_y_continuous(breaks = seq(0, 350, by = 50)) +
  geom_vline(xintercept = 1923) +
  annotate("text", x = 1928, y = 50, label = "Disney Brothers \nCartoon Studio") +
  geom_vline(xintercept = 1983) +
  annotate("text", x = 1984, y = 100, label = "Disney Channel \n on TV") +
  geom_vline(xintercept = 2019) +
  annotate("text", x = 2016, y = 300, label = "Disney+ streaming \n service") +
  labs(title = "Amount of Movies and TV Shows Released", 
       x = "Release Year",
       y = "Amount") +
  theme_bw() +
  theme(legend.title = element_blank()) +
  scale_fill_manual(values = c("darkseagreen3", "darkslategray2"))
```

#### Release Years by Program Type---Summary

Create a quantitative summary of `release_year` by program `type` that shows the oldest release year, the median release year, and the newest release year for each group of programs (i.e., movies and TV shows).

```{r summarystats}
movie_release <- tidy_disney_data %>%
  group_by(type) %>%
  select(release_year) %>%
  filter(type == "Movie")
summary(movie_release)

show_release <- tidy_disney_data %>%
  group_by(type) %>%
  select(release_year) %>%
  filter(type == "TV Show")
summary(show_release)
```

### Analysis 2: Disney Content & Ratings

**This analysis uses Dataset #1.**

Create a horizontal barplot to show the frequencies of the various `rating` categories for programs on Disney+ (i.e., use `y =` rather than `x =` for the variable you are plotting). **You will need to remove a couple `NA` values before plotting** Facet by `type` so that you have side-by-side plots for the two different program types. Add an informative title and axis labels, as well as at least two other types of useful formatting. 

```{r ratingsbarplot}
tidy_disney_data %>%
  na.omit() %>%
  ggplot(aes(y = rating)) +
    geom_bar() +
    facet_grid("type") +
  labs(title = "Number of Ratings on Disney+ Movies and Shows",
       y = "Rating",
       x = "Count") +
  theme_bw() +
  scale_x_continuous(breaks = seq(0, 275, by = 25))
```

### Analysis 3: Content on Different Services

**This analysis uses Dataset #2.**

#### Quantity of Content by Service

Create an appropriate barplot to show the number of TV shows offered by the four streaming services. Include an informative title and axis labels. Add at least two other types of formatting to enhance the effectiveness and/or appearance of your plot (e.g., themes, colors, etc).

```{r servicequantity}
tidy_streaming_data %>%
  ggplot(aes(y = YesNo, fill = Service)) +
  geom_bar() +
  facet_grid("Service") +
  theme_bw() +
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        legend.position = "none") +
  labs(title = "Amount of TV Shows on Different Streaming Services",
       x = "Count",
       y = "Streaming Services") +
  scale_x_continuous(breaks = seq(0, 2000, by = 200)) +
  scale_fill_manual(values = c("blue2", "chartreuse3", "brown3", "deepskyblue"))
```

#### Quality of Content by Service

Create an appropriate plot to compare either IMDB or Rotten Tomatoes ratings for content on the streaming services. Include an informative title and axis labels. Add at least two other types of formatting to enhance your plot (e.g., themes, colors, etc).

```{r serviceratings}
tidy_streaming_data %>%
  ggplot(aes(x = Service, y = rt_score, fill = Service)) +
  geom_boxplot(outlier.size = 0.7) +
  theme_bw() +
  labs(title = "Distribution of Rotten Tomato Scores for Streaming Services",
       x = "Streaming Services",
       y = "Rotten Tomato Score") +
  theme(legend.position = "none") + 
  scale_fill_manual(values = c("blue2", "chartreuse3", "brown3", "deepskyblue"))
```

### Analysis 4: Choose Your Own Plot!

**This plot uses either dataset.**

Create one other plot of your choice using either dataset to explore a question of interest. It should not be too similar to any of the plots you made above. You are welcome to perform additional manipulations on the data, if needed. Add an appropriate title and axis labels, as well as any other necessary formatting to make your plot clear and effective.

```{r choiceplot}
tidy_streaming_data %>%
  na.omit() %>%
  filter(Service == "Netflix") %>%
  ggplot(aes(x = Age)) +
  geom_bar() +
  labs(title = "Distribution of Age Ratings on Netflix",
       x = "Age Rating",
       y = "Count") +
  theme_bw() +
  scale_x_discrete(limits = c("7+", "13+", "16+", "18+", "all")) +
  scale_y_continuous(breaks = seq(0, 500, by = 50))
```

***

## Questions

**Question 1:** Based on your plots, make five informational statements or comparisons regarding the Disney+ streaming service.

**ANSWER**

1) Disney+ has the least amount of shows compared to the other three streaming services. 

2) The amount of TV Shows released by Disney is left skewed meaning a majority of TV shows were released more recently.

3) Most Disney movies are rated G while not many movies are rated TV-Y. 

4) Most Disney TV shows are rated TV-PG and no shows are rated PG-13.

5) Disney had an increase in TV shows released after the creation of Disney Channel in 1983

**Question 2:** What other data would you like to have, or which existing variables would you like to see transformed, if you were going to do further explorations or visualizations? Give at least two examples.

**ANSWER**

We would like to see the streaming service HBOMax added to the streaming data. In the Disney+ dataset, we would like to add a variable about the number of views a TV show or movie has to see how popular the content is.   

**Question 3:** Explain the rationale behind the choices you made with regard to plot type, formatting, and so on, when you created Analysis 4. What motivated your decisions with regard to this plot? 

**ANSWER**

We wanted to see the rating distribution of Netflix shows because the age of people that use Netflix is very broad, so looking at the age rating of the content they are creating can show what age demographics they may target the most. A bar plot was the plot we chose because this shows the count of each age rating, so it would be easy to see which rating had the most and least. 

***
```{r}
sessionInfo()
```
